\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-05-27 scratch/notes_john/JSS-article/JSS-article-3.Rmd
simply use of caret in example in JSS draft paper

The \pkg{caret} package doesn't directly support generalized linear models fit by \code{glm()}, but it can be made to do so by defining a custom "method" for the binary logistic regression fit to the \code{attrition} data:
```{r rsample-example-using-caret, cache=TRUE}
library("caret", quietly=TRUE, warn.conflicts=FALSE)

glm_mod <- getModelInfo("glm", regex = FALSE)[[1]]
glm_mod$fit <- function(x, y, wts, param, lev, last, classProbs, ...) {
  x$Attrition <- y
  glm(Attrition ~ JobSatisfaction + Gender + MonthlyIncome,
      family = binomial, data = x)
}

mod <- train(x = attrition[, c("JobSatisfaction", "Gender",
                               "MonthlyIncome")],
             y = attrition$Attrition,
             method = glm_mod,
             trControl = trainControl(method="LOOCV"))
mod # reports complement of Bayes-rule criterion
```

@@

microbenchmark::microbenchmark(
  rsample={
    rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
    rs_obj$accuracy <- map_dbl(rs_obj$results,
                               function(x) mean(x$correct))
  },
  caret=train(x = attrition[, c("JobSatisfaction", "Gender",
                                "MonthlyIncome")],
              y = attrition$Attrition,
              method = glm_mod,
              trControl = trainControl(method="LOOCV")),
  direct={
    n <- nrow(attrition)
    yhat <- numeric(n)
    for (i in 1:n){
      m <- update(mod.attrition, data=attrition[-i, ])
      yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
    }
    mean(((attrition$Attrition == "Yes") - round(yhat))^2)
  },
  cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
  cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
             method="Woodbury"),
  cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
                  method="hatvalues"),
  times=10
)
```
Computing time for \pkg{rsample} and \pkg{caret} for this problem is similar to direct computation of LOO CV and to \code{cv()} with \code{method="exact"}, but more than an order of magnitude slower than \code{cv()} with \code{method="Woodbury"}, and more than three orders of magnitude slower then \code{cv()} with \code{method="hatvalues"}.

>>>>

The \pkg{caret} package supports generalized linear models fit by \code{glm()} and can straightforwardly perform LOO CV:
```{r rsample-example-using-caret, cache=TRUE}
library("caret", quietly=TRUE, warn.conflicts=FALSE)
# reports complement of Bayes-rule criterion:
train(x = attrition[, c("JobSatisfaction", "Gender",
                               "MonthlyIncome")],
             y = attrition$Attrition,
             method = "glm",
             trControl = trainControl(method="LOOCV"))
```

@@

microbenchmark::microbenchmark(
  rsample={
    rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
    rs_obj$accuracy <- map_dbl(rs_obj$results,
                               function(x) mean(x$correct))
  },
  caret=train(x = attrition[, c("JobSatisfaction", "Gender",
                                "MonthlyIncome")],
              y = attrition$Attrition,
              method = "glm",
              trControl = trainControl(method="LOOCV")),
  direct={
    n <- nrow(attrition)
    yhat <- numeric(n)
    for (i in 1:n){
      m <- update(mod.attrition, data=attrition[-i, ])
      yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
    }
    mean(((attrition$Attrition == "Yes") - round(yhat))^2)
  },
  cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
  cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
             method="Woodbury"),
  cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
                  method="hatvalues"),
  times=10
)
```
Computing time for \pkg{rsample} and \pkg{caret} for this problem is similar to direct computation of LOO CV and to \code{cv()} with \code{method="exact"}, but more than an order of magnitude slower than \code{cv()} with \code{method="Woodbury"}, and more than three orders of magnitude slower then \code{cv()} with \code{method="hatvalues"}.

### 2024-05-27 scratch/notes_john/JSS-article/JSS-article-3.tex
simply use of caret in example in JSS draft paper

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 799417
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.20609
bias-adjusted cross-validation criterion = 19.19442
95% CI for bias-adjusted CV criterion = (15.73743, 22.6514)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x1118adc78>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

@@

The \pkg{caret} package doesn't directly support generalized linear
models fit by \code{glm()}, but it can be made to do so by defining a
custom ``method'' for the binary logistic regression fit to the
\code{attrition} data:

\begin{CodeChunk}
\begin{CodeInput}
R> library("caret", quietly=TRUE, warn.conflicts=FALSE)
R> 
R> glm_mod <- getModelInfo("glm", regex = FALSE)[[1]]
R> glm_mod$fit <- function(x, y, wts, param, lev, last, classProbs, ...) {
+   x$Attrition <- y
+   glm(Attrition ~ JobSatisfaction + Gender + MonthlyIncome,
+       family = binomial, data = x)
+ }
R> 
R> mod <- train(x = attrition[, c("JobSatisfaction", "Gender",
+                                "MonthlyIncome")],
+              y = attrition$Attrition,
+              method = glm_mod,
+              trControl = trainControl(method="LOOCV"))
R> mod # reports complement of Bayes-rule criterion
\end{CodeInput}
\begin{CodeOutput}
Generalized Linear Model

@@

\begin{CodeChunk}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury")$`CV crit`,
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="hatvalues")$`CV crit`
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   rsample={
+     rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
+     rs_obj$accuracy <- map_dbl(rs_obj$results,
+                                function(x) mean(x$correct))
+   },
+   caret=train(x = attrition[, c("JobSatisfaction", "Gender",
+                                 "MonthlyIncome")],
+               y = attrition$Attrition,
+               method = glm_mod,
+               trControl = trainControl(method="LOOCV")),
+   direct={
+     n <- nrow(attrition)
+     yhat <- numeric(n)
+     for (i in 1:n){
+       m <- update(mod.attrition, data=attrition[-i, ])
+       yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
+     }
+     mean(((attrition$Attrition == "Yes") - round(yhat))^2)
+   },
+   cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
+   cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury"),
+   cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
+                   method="hatvalues"),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(rsample = {: less accurate nanosecond
times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: milliseconds
         expr         min          lq        mean      median          uq
      rsample 5853.035606 5948.291603 6163.101903 6121.042951 6144.223100
        caret 6804.763276 6951.219335 7097.717763 7061.439656 7255.566640
       direct 5310.127259 5350.848336 5554.513023 5484.144809 5690.480487
     cv.exact 4751.498733 4797.980556 4879.481352 4870.303039 4927.941413
      cv.wood  101.606159  103.891294  108.285899  108.577430  112.075263
 cv.hatvalues    1.206999    1.471613    1.514548    1.554679    1.614416
         max neval
 6981.899881    10
 7541.571242    10
 6081.412412    10
 5118.834994    10
  116.574603    10
    1.759433    10
\end{CodeOutput}
\end{CodeChunk}

>>>>

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 45649
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.13181
bias-adjusted cross-validation criterion = 19.12404
95% CI for bias-adjusted CV criterion = (15.69034, 22.55774)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x145d6f3a0>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

@@

The \pkg{caret} package supports generalized linear models fit by
\code{glm()} and can straightforwardly perform LOO CV:

\begin{CodeChunk}
\begin{CodeInput}
R> library("caret", quietly=TRUE, warn.conflicts=FALSE)
R> # reports complement of Bayes-rule criterion:
R> train(x = attrition[, c("JobSatisfaction", "Gender",
+                                "MonthlyIncome")],
+              y = attrition$Attrition,
+              method = "glm",
+              trControl = trainControl(method="LOOCV"))
\end{CodeInput}
\begin{CodeOutput}
Generalized Linear Model

@@

\begin{CodeChunk}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury")$`CV crit`,
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="hatvalues")$`CV crit`
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   rsample={
+     rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
+     rs_obj$accuracy <- map_dbl(rs_obj$results,
+                                function(x) mean(x$correct))
+   },
+   caret=train(x = attrition[, c("JobSatisfaction", "Gender",
+                                 "MonthlyIncome")],
+               y = attrition$Attrition,
+               method = "glm",
+               trControl = trainControl(method="LOOCV")),
+   direct={
+     n <- nrow(attrition)
+     yhat <- numeric(n)
+     for (i in 1:n){
+       m <- update(mod.attrition, data=attrition[-i, ])
+       yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
+     }
+     mean(((attrition$Attrition == "Yes") - round(yhat))^2)
+   },
+   cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
+   cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury"),
+   cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
+                   method="hatvalues"),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(rsample = {: less accurate nanosecond
times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: milliseconds
         expr         min          lq        mean      median          uq
      rsample 5741.137299 5867.248666 6155.761099 6095.937708 6218.031915
        caret 6805.241869 6931.865408 7135.256404 7052.951569 7377.336681
       direct 5262.786773 5312.612884 5455.994681 5388.428711 5475.565416
     cv.exact 4740.642384 4771.993936 4870.961647 4847.941434 4981.140799
      cv.wood  104.161361  106.036578  112.172855  113.107500  115.602534
 cv.hatvalues    1.262554    1.462716    1.497295    1.520198    1.569685
         max neval
 7162.266671    10
 7648.302729    10
 6045.941344    10
 5054.087753    10
  120.361527    10
    1.666896    10
\end{CodeOutput}
\end{CodeChunk}
